Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
getwd()
pf <- read.delim('pseudo_facebook.tsv')
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
install.packages('ggthemes', dependencies = TRUE)
library(ggthemes)
theme_set(theme_minimal(24))
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
# facet_grid is more useful when analysing two or more variables
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~dob_month, ncol = 3)

# equivalent ggplot syntax
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:
Outliers are top-coded to be monitored or replaced by a normal value.
***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(data = pf, x = friend_count)
# Equivalent ggplot syntax for solution:

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram()
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(data = pf, x = friend_count, xlim = c(0,1000))
# or
qplot(data = pf, x = friend_count) + scale_x_continuous(limits = c(0, 1000))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
# binwidth to set up the width of the bar; breaks to set up the lable steps on the x axis
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
# Equivalent ggplot syntax:

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
# This code below will only remove rows if gender = na
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)

# This coded below will remove the row if any of the value = na
qplot(x = friend_count, data = na.omit(pf), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:women  

#### What's the difference between the median friend count for women and men?
Response:22

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}
#The parameter color determines the color outline of objects in a plot.

#The parameter fill determines the color of the area inside objects in a plot.

#You might notice how the color black and the hex code color of #099DD9 (a shade of blue) are wrapped inside of I(). The I() functions stand for 'as is' and tells qplot to use them as colors.
qplot(data = pf, x = tenure, binwidth = 30, color = I('black'), fill = I('#099DD9'))
# Equivalent ggplot syntax for plots:

ggplot(aes(x = tenure), data = pf) +
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')

ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(data = pf, x = tenure/365, binwidth = .25, color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7))
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(data = pf, x = tenure/365, 
      xlab = 'Number of years using Facebook',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), lim = c(0, 7))

# Equivalent ggplot syntax:

ggplot(aes(x = tenure / 365), data = pf) +
  geom_histogram(color = 'black', fill = '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```

***

### User Ages
Notes:

```{r User Ages}


qplot(data = pf, x = age, binwidth = 1, xlab = 'Age of Facebook users', ylab = 'Number of users in sample', color = I('black'), fill = I('#467F18'))


qplot(data = pf, x = age, binwidth = 1, xlab = 'Age of Facebook users', ylab = 'Number of users in sample', color = I('black'), fill = I('#467F18')) +
  scale_x_continuous(breaks = seq(0, 100, 25), limits = c(1, 100))
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***
```{r Transforming Data}
qplot(x = friend_count, data = pf)
summary(pf$friend_count)
summary(log10(pf$friend_count + 1))
summary(sqrt(pf$friend_count))

install.packages('gridExtra')
library(gridExtra)
# By using the below code, the x axis lable is different as the actual friend count
p1 = qplot(x = friend_count, data = pf)
p2 = qplot(x = log10(friend_count + 1), data = pf) 
p3 = qplot(x = friend_count, data = pf) + scale_x_log10()
p4 = qplot(x = sqrt(friend_count), data = pf)
grid.arrange(p1,p2,p3, p4)
```

### Transforming Data Alternate Solution
### Use scales!
### By using the layers of scales, the x axis lable is the actual friend count.
```{r Transforming Data scales}
p1 = ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 = p1 + scale_x_log10()
p3 = p1 + scale_x_sqrt()
grid.arrange(p1, p2, p3)
```
### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)), 
      geom = 'freqpoly', binwidth = 10, color = gender,
      xlab = 'Friend Count',
      ylab = 'Proportion of users with that friend count') +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) 

# Equivalent ggplot syntax:

ggplot(aes(x = friend_count, y = ..count../sum(..count..)),
       data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
by(pf$www_likes, pf$gender, sum)
qplot(x = www_likes, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)), 
      geom = 'freqpoly', color = gender,
      xlab = 'Friend Count',
      ylab = 'Proportion of users with that friend count') + scale_x_continuous() + scale_x_log10()
```


***

### Box Plots
Notes:

```{r Box Plots}
# remember to use continuous variable as y axis and categorical variable as x axis for plotting box plot
qplot(geom = 'boxplot', data = subset(pf, !is.na(gender)), x = gender, y = friend_count, ylim = c(0, 1000))
# or 
qplot(geom = 'boxplot', data = subset(pf, !is.na(gender)), x = gender, y = friend_count) + scale_y_continuous(limits = c(0, 1000))
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
# remember to use continuous variable as y axis and categorical variable as x axis for plotting box plot
qplot(geom = 'boxplot', data = subset(pf, !is.na(gender)), x = gender, y = friend_count, ylim = c(0, 1000))
# or 
qplot(geom = 'boxplot', data = subset(pf, !is.na(gender)), x = gender, y = friend_count) + coord_cartesian(ylim = c(0,1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
# remember to use continuous variable as y axis and categorical variable as x axis for plotting box plot
qplot(geom = 'boxplot', data = subset(pf, !is.na(gender)), x = gender, y = friend_count) + coord_cartesian(ylim = c(0, 250))

by(pf$friend_count, pf$gender, summary)

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
names(pf)
qplot(geom = 'boxplot', data = subset(pf, !is.na(gender)), x = gender, y = friendships_initiated) + coord_cartesian(ylim = c(0, 200))
by(pf$friendships_initiated, pf$gender, summary)
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes>0)
mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
levels(pf$mobile_check_in)
summary(pf$mobile_check_in)
sum(pf$mobile_check_in == 1) / length(pf$mobile_check_in)
```

Response:

***

### Analyzing One Variable
Reflection:
1. Taking close look at the individual variables in the dataset;
2. Understanding the types of values
3. What the distribution looks like
4. whether there are missing values or outliers
5. Histogram, boxplot, frequency polygon for visualising and understanding data
6. making adjustment to these plots e.g change the limit of axises, adjusting the binwidth of the histogram, transforming variable log/sqrt/binaries to uncover hidden patterns of the data
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!