Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
setwd('/Users/arielma/data-analysis/lesson5_EDA/5_explore_two_variables')
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
qplot(x = age, y = friend_count, data = pf)
# or
qplot(age, friend_count, data = pf)

# or in ggplot syntax - can make more complex plot
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point() + 
  xlim(13, 90)

summary(pf$age)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point() + 
  xlim(13, 90)
```

***

### Overplotting
Notes: Makes it hard to figure out how many points are in each region

```{r Overplotting}
# alpha = 1/20 means it will take 20 points to be equavelent to one of the black solid dots
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20) + 
  xlim(13, 90)
# ages are continuous variable but in the data set we only have integer values. So we see the perfect lined up columns which is not a true reflection of age. These columns should fee intuitively right, and we want to make sure we can see more of the points. So using jitter we can add some noise (positive or negative noise to each of our points) in each age so we get a clear picture between age and friend_count.

ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_jitter(alpha = 1/20) + 
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20) + 
  xlim(13, 90) +
  coord_trans(y = 'sqrt')

# to use jitter together with coord_trans, we have to use more elaborate syntax to specify that we only want to jitter the ages. We also need to be careful since some people have a friend count of 0. If we add noise to 9 friend counts, we might end up with negative numbers for some of our friend counts and those square roots would be imaginary.
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13, 90) +
  coord_trans(y = 'sqrt')

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
names(pf)
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  coord_trans(y = 'sqrt')
  

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
# dplyr package lets us split up a data frame and apply a function to some parts of the data.
install.packages('dplyr')
library(dplyr)
# Some common functions can be used in dplyr are filer(), group_by(), mutate(), arrange()
# n = n() is the number of users in each group. This n() function can only be used for summarise() and reports how many people are really in each group.
age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)
```

Create your plot!

```{r Conditional Means Plot}
# using %>% to chain functions so that we dont need to save the results in variables and pass then into the next function
pf.fc_by_age = pf %>% 
  group_by(age) %>% 
  summarise(friend_count_mean = mean(friend_count),
                                   friend_count_median = median(friend_count),
                                   n = n()) %>%
  arrange(age)
head(pf.fc_by_age)

ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20, 
             position = position_jitter(h = 0),
             color = 'orange') + 
  geom_line(stat = 'summary', fun.y = mean, color = 'red') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .1), linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .9), linetype = 2, color = 'green') +
  geom_line(stat = 'summary', fun.y = median) +
  coord_cartesian(xlim = c(13, 70), ylim = c(0,1000))
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
?cor.test
cor.test(pf$age, pf$friend_count, method = 'pearson')
# the correlation coefficient of 0.0274. THis indicates that there is no meaningful relationship between the two variables.
# A good rule of thumb is that a correlation greater than 0.3 or less than -0.3 is meaningful but small; around 0.5 is moderate and 0.7 or greater is large 
# or
with(pf, cor.test(age, friend_count, method = 'pearson'))
# the with function lets us evaluate an R expression in an environment constructed from the data
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
# if ignoring the method argument, cor.test will use pearson product moment correlation as default.
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
# method = 'lm' meaning using linear model
ggplot(data = pf, aes(www_likes_received, likes_received)) + geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
cor.test(pf$www_likes_received, pf$likes_received)

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:
One of the assumptions of regression is the measured variables are independent of each other. So if any two are too highly related with each other, it will be really difficult to tell which ones are actually driving the phenomenon. And so it's important to measure the correlation between your variables first, often because it will help you determine which ones you dont want to throw in together, and it might help you decide which ones you actually want to keep.
***

### More Caution with Correlation
Notes:
Correlation can help us decide which variables are related. But even correlation coefficients can be deceptive if you are not careful. Plotting the data is the best way to help you understand it and it can lead you to key insights.
```{r More Caution With Correlation}
install.packages('alr3')

install.packages('car')
install.packages('carData')
install.packages('readxl')
library(car)
library(alr3)
library(ggplot2)
data(Mitchell)
?Mitchell
```

Create your plot!

```{r Temp vs Month}

ggplot(data = Mitchell, aes(x = Month, y = Temp)) + geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
range(Mitchell$Month)
ggplot(data = Mitchell, aes(x = Month, y = Temp)) + geom_point() + 
  scale_x_continuous(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:
This is an example of how it's so important to get perspective on your data. You want to make sure you put your data in context. Another important point to make is that the proportion and scale of the graphics do matter. 
The nature of the data should suggest the shape of the graphic. Otherwise, you should tend to have a graphic that's about 50% wider than it is tall.
```{r}
ggplot(aes(x=(Month%%12),y=Temp), data=Mitchell)+
  geom_point()
```

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + geom_line()
pf$age_with_months = pf$age + (12 - pf$dob_month)/12
```

***

### Age with Months Means

```{r Age with Months Means}
library(dplyr)

pf.fc_by_age_months = pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(as.numeric(friend_count)),
            n = n()) %>%
  arrange(age_with_months)

names(pf)
```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
    ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) + geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean), data = subset(pf.fc_by_age, age < 71)) + geom_line() + geom_smooth()
p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) + geom_line() + geom_smooth()

library(gridExtra)
grid.arrange(p2, p1, ncol = 1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

